---
title: "DRY coding"
author: "Data Science in Biomedicine - Week 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
## Introduction

We'll be using an in-built data set from the MASS package. But, because the MASS package can be a bit problematic to install, I've provided the data for you. It's in .rds format, so we load it in using the `read_rds()` function.

For more info about the dataset, head to https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/survey.html

```{r}
library(tidyverse)
library(knitr)
survey <- read_rds("./data/survey.rds")
```

## Exercise 1
Explore the data - what variables are in there, what variable types, how many variables and observations?
```{r}
str(survey)
```


## Exercise 2
Round every numeric variable to 0 decimal places.

```{r}
survey_round <- survey %>%
  mutate(across(where(is.numeric), round))
```


## Exercise 3
Using the rounded dataframe you created in Exercise 2, calculate the mean of each numeric variable in the dataset. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey_round_mean <- survey_round %>%
  select(where(is.numeric)) %>%
  map_df(~mean(.x,na.rm = TRUE)) %>%
  kable(
    col.names = c("Writing Hand", "Non-writing Hand", "Pulse","Height", "Age"),
    align = "ccccc"
  )
survey_round_mean
```


## Exercise 4
Now make exactly the same table, using the pivot functions. Again, use kable() to insert a nicely formatted table into your R Markdown document.
```{r}
survey_round_mean_pivot <- survey_round %>% 
  select (where(is.numeric)) %>%
  pivot_longer(
    cols = everything(),
  ) %>%
  group_by(name) %>%
  summarise(mean = mean(value, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = name,
    values_from = mean
  ) %>%
  kable()
survey_round_mean_pivot
```

## Exercise 5
Students had initially expressed their height in either imperial or metric units, but someone converted everything into centimeters. Create a new variable, Height_original, which will show height in centimeters for students who expressed their height in metric units, and height in inches for students who prefer imperial units. 

NOTE: This one is a bit of a challenge. You may want to check out the `case_when()` function.

```{r}
survey_height <- survey_round %>%
mutate(Height_original = case_when(
  M.I == "Metric"~Height,
  M.I == "Imperial"~Height/2.54
))
```


## Exercise 6
Create a table that shows the mean and standard deviation of height - in centimeters for students who used metric units, and in inches for students who used imperial units. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey_height_mean_sd <- survey_height %>%
  filter(!is.na(M.I)) %>%
  group_by(M.I) %>%
  summarise(sd = sd(Height_original), mean = mean(Height_original))
  
```


## Exercise 7
Create a contingency table to see the relationship between exercise and smoking. Pivot your table to make it easier to read, and then use kable() to insert a nicely formatted table into your R Markdown document.

```{r}

```


## Exercise 8
Create a grid of Pulse histograms - one histogram for each combination of Sex and Exercise.

```{r}

```


## Exercise 9
Create a new variable, Hnd.diff, that shows the difference in span between the writing and non-writing hand. Create a set of boxplots of this variable, with a separate boxplot for those who write with their right hand, and another one for those who write with their left hand.

```{r}

```

